---
title: "Homework 1 - readable and efficient R code"
author: "Tian Zhao"
date: "`r format(Sys.time(), '%m/%d/%Y')`"
output: html_document
---

```{r}
library(microbenchmark)
```


# Question 1 - "function-alize" this code

Read over the code below and perform the following:

* Wrap it into a function `foobar0` which has arguments `x` and `z`
  and which returns the vector `x` at the end of the following code.
* Rewrite this into a function `foobar` which is easier to read, by
  reducing repetitive code. E.g. `foobar` might call a function to
  check the input, and another function to perform the three lines of
  computation. 
* Check that the two versions produce the same output using the
  function `all.equal`.

```{r}
set.seed(1)
x <- rnorm(100)
z <- rnorm(100)
if (sum(x >= .001) < 1) {
  stop("step 1 requires 1 observation(s) with value >= .001")
}
fit <- lm(x ~ z)
r <- fit$residuals
x <- sin(r) + .01
if (sum(x >= .002) < 2) {
  stop("step 2 requires 2 observation(s) with value >= .002")
}
fit <- lm(x ~ z)
r <- fit$residuals
x <- 2 * sin(r) + .02
if (sum(x >= .003) < 3) {
  stop("step 3 requires 3 observation(s) with value >= .003")
}
fit <- lm(x ~ z)
r <- fit$residuals
x <- 3 * sin(r) + .03
if (sum(x >= .004) < 4) {
  stop("step 4 requires 4 observation(s) with value >= .004")
}
fit <- lm(x ~ z)
r <- fit$residuals
x <- 4 * sin(r) + .04
x
```

**a**

```{r}
foobar0 <- function(x,z){
  if (sum(x>=0.001)<1){
    stop("step 1 requires 1 observation(s) with value >= .001")
  }
  fit <- lm(x~z)
  r <- fit$residuals
  x <- sin(r)+0.01
  if (sum(x>=0.002)<2){
    stop("step 2 requires 2 observation(s) with value >= .002")
  }
  fit <- lm(x~z)
  r <- fit$residuals
  x <- 2*sin(r)+0.02
  if (sum(x>=0.003)<3){
    stop("step 3 requires 3 observation(s) with value >= .003")
  }
  fit <- lm(x~z)
  r <- fit$residuals
  x <- 3*sin(r)+0.03
  if (sum(x>=0.004)<4){
    stop("step 4 requires 4 observation(s) with value >= .004")
  }
  fit <- lm(x~z)
  r <- fit$residuals
  x <- 4*sin(r)+0.04
  return(x)
}
```

**b**

```{r}
get_x <- function(x,z,i){
  if (sum(x >= 0.001*i)<i){
    stop(paste("step",i,"requires",i,"observation(s)","with value >=",i*0.001))
  }
  fit <- lm(x~z)
  r <- fit$residuals
  x <- i*sin(r)+0.01*i
  return(x)
}
```

```{r}
foobar <- function(x,z){
  for (i in 1:4){
    x <- get_x(x,z,i)
  }
  return(x)
}
```

**c**

```{r}
set.seed(1)
x <- rnorm(100)
z <- rnorm(100)
all.equal(foobar(x,z),foobar0(x,z))
```


# Question 2 - vectorize this code and benchmark

* Take the following function `f0` and rewrite it as a function `f`,
  which is faster and easier to read, by removing the loop of `i` from
  1 to `m`.
* Benchmark `f` and `f0` using `microbenchmark`. How much faster is `f`?

```{r}
n <- 30
p <- 50
p2 <- 25
m <- 1000
set.seed(1)
x <- matrix(rnorm(n*p),nrow=n,ncol=p)
b <- matrix(rnorm(m*p),nrow=m,ncol=p)
a <- matrix(rnorm(m*p2),nrow=m,ncol=p2)
f0 <- function(x,b,a) {
  out <- numeric(0)
  for (i in seq_len(m)) {
    bb <- b[i,]
    aa <- a[i,]
    out <- c(out, sum(x %*% bb) + sum(aa))
  }
  out
}
```

```{r}
f <- function(x,b,a){
  return(apply(a,1,sum)+apply(x%*%t(b),2,sum))
}
```

```{r}
all.equal(f(x,b,a),f0(x,b,a))
```

```{r}
microbenchmark(f(x,b,a),f0(x,b,a))
```

We can see that the modified function is on average 50% faster than the old version.

# Question 3 - build a faster t-test

* Rewrite the following function `getT0` which computes `m`
  two-sample t-tests (equal variance) between two groups as a function
  `getT`, which is faster by using vectorized operations over the `m`
  sets of observations. (There are functions in R packages, such as
  `genefilter::rowttests` which will quickly perform this operation,
  but I want you to write your own function using simple R functions
  like `rowSums`, etc.)
* Benchmark `getT` and `getT0`. How much faster is `getT`?

```{r}
m <- 400
n <- 50
little.n <- n/2
set.seed(1)
x <- matrix(rnorm(m*n),nrow=m,ncol=n)
f <- gl(2,little.n)
getT0 <- function(x, f) {
  ts <- sapply(seq_len(m), function(i) t.test(x[i,] ~ f, var.equal=TRUE)$statistic)
  unname(ts)
}
```

```{r}
rowvars <- function(x){
  n <- dim(x)[2]
  return(rowSums((x-rowMeans(x))^2)*(1/(n-1)))
}
```


```{r}
getT <- function(x,f){
  x1 <- x[,f==unique(f)[1]]
  x2 <- x[,f==unique(f)[2]]
  n1 <- dim(x1)[2]
  n2 <- dim(x2)[2]
  t <- (rowMeans(x1)-rowMeans(x2))/sqrt(((n1-1)*rowvars(x1)+(n2-1)*rowvars(x2))/(n1+n2-2)*(1/n1+1/n2))
  return(t)
}
```



```{r}
all.equal(getT(x,f),getT0(x,f))
```

```{r}
microbenchmark(getT(x,f),getT0(x,f))
```




